Appendix to ``A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol'' published in the Journal of Computational and Graphical Statistics

### 2024-03-24 JCGS_response_to_reviews/response_to_review_R1.Rmd

---
title: "Response to Reviews: JCGS-23-139.R1"
output: pdf_document
date: "`r Sys.Date()`"
---

@@

Thank you for the positive feedback.

# Reviewer 2

*I appreciate the effort provided by the authors for both my comments and the other reviewer’s comments. The revision is much improved. I have a few remaining comments.*

We appreciate the reviewer's acknowledgment of our efforts in addressing the comments and improving the manuscript. Thank you for your remaining comments, which we have carefully considered and addressed in the revised version.

@@

*1. I still believe the paper is too long. Starting with Section 4, I believe the paper starts to become a little too cumbersome. I suggest the authors look for ways to condense down Sections 4 and 5 a bit as I think these are the biggest contributors to the length.*

Thank you for the suggestion to condense Sections 4 and 5. We agree that the length of these sections contributed to the overall length of the paper. In the revised version, we have streamlined these sections, reducing the main text by 4 pages. Some of the details have been moved to the appendix to improve readability while retaining the necessary information.

@@

*2. In the Figure 1 caption, clarify what is meant by the "at position $2^2 + 2$".*

Thank you for seeking clarification on the caption of Figure 1. The expression $2^2 + 2$ was intended to introduce a brief pause before revealing the data plot, allowing the reader time to examine the lineup first. However, we agree that this notation may be unclear. In the revised version, we have replaced it with the numerical value $6$, which represents the position where the data plot is presented.

@@

*3. In the footnote on the bottom of page6, replace "didn't" and "it's" by "did not" and "it is", respectively.*

Thank you for pointing out the use of contractions in the footnote on page 6. We have replaced "didn't" with "did not" and "it's" with "it is" as suggested. Additionally, we have carefully reviewed the manuscript and addressed any other inappropriate use of contractions to ensure appropriate academic writing style throughout the text.

@@

*4. In the last sentence of Section 2.2, the final segment "statistical and practical statistics" sounds awkward because "statistical" should be serving as an adjective to another word, and the segment as written makes it sound like it is also an adjective to "statistics". The authors might consider replacing "statistical" with "classic inference" as the paper is addressing the discrepancies between inferential results and what may be practically done.*

Thank you for the insightful feedback on the language used in the last sentence of Section 2.2. We appreciate you pointing out the awkward phrasing of "statistical and practical statistics." As you suggested, the use of "statistical" as an adjective modifying "statistics" is redundant and may cause confusion. In the revised version, we have replaced that phrase with "statistical significance and practical significance" to better reflect the contrast between inferential results and practical implications, which is the focus of that discussion.

@@

*5. Page 21 onward, there are many instances where the figure references did not compile correctly. These need to be corrected.*

Thank you for bringing to our attention the issue with the figure references not compiling correctly from page 21 onward. We have carefully reviewed and recompiled the manuscript, ensuring that all figure references are now accurately represented and correctly linked.

### 2024-03-24 appendix.Rmd
R2 submitted

---
title: |
  Appendix: A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
  - name: Dianne Cook
    affil: a
    email: dicook@monash.edu
  - name: Emi Tanaka
    affil: a
    email: emi.tanaka@monash.edu
  - name: Susan VanderPlas
    affil: b
    email: susan.vanderplas@unl.edu
affiliation:
  - num: a
    address: |
      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
  - num: b
    address: |
      Department of Statistics, University of Nebraska, Lincoln, Nebraska, USA
bibliography: ref.bib
header-includes: |
  \usepackage{lscape}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
  \doublespacing
output: rticles::tf_article
editor_options: 
  chunk_output_type: console
---

@@

# Additional details of experimental design

## Non-linearity

## Heteroskedasticity

Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively. A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern. 
The parameter $b$ controls the variation in the standard deviation of the error across the support of the predictor. A larger value of $b$ generates a larger ratio between it's smallest and highest values, making the visual pattern more obvious. 
The sample size $n$ sharpens (or blurs) a pattern when it is larger (or smaller).

<!-- Figures \ref{fig:different-sigma}, \ref{fig:different-b} and \ref{fig:different-n} demonstrate the impact of these parameters on signal strength.  -->

### Lineup allocation to participants

From Table \ref{tab:model-factor-table} it can be computed that there are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ combinations of parameter values for the non-linearity model and heteroskedasticity models respectively. Three replications for each combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups, respectively.

Each lineup needs to be evaluated by at least five participants. From previous work, and additional pilot studies for this experiment, we decided to record evaluations from 20 lineups for each participant. Two of the 20 lineups with clear visual patterns were used as attention checks, to ensure quality data. Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ participants were needed to cover the experimental design for the data collection periods I and II, respectively. The factor levels and range of difficulty was assigned relatively equally among participants. The Appendix contains graphical summaries of these assignments for each subject.

Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$ (see Section A.1 in the Appendix), and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results. To construct a Rorschach lineup, the data is generated from a model with zero effect size, while the null data are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}. This procedure differs from that of the canonical Rorschach lineup, where all 20 plots are generated directly from the null model, so the method suggested in @vanderplas2021statistical for typical lineups containing a data plot is used to estimate $\alpha$. (The Appendix contains a sensitivity analysis on the effect of uncertainty in the estimation of $\alpha$.) The $3 \times 4 = 12$ treatment levels of the common factors, replicated three times results in 36 lineups. And 6 more evaluations on the 279 lineups with uniform fitted value distribution, results in at least $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants needed.

### Collecting results

Participants for all three data collection periods were recruited from the Prolific crowd-sourcing platform [@palan2018prolific]. Pre-screening procedures were applied during the recruitment: participants were required to be fluent in English, with $98\%$ minimum approval rate and at least 10 submissions in other studies.

In each lineup evaluation, the participant was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others. If there is no noticeable difference between plots in a lineup, participants had the option to select zero plots without the need to provide a reason. During the process of recording the responses, a zero selection was considered to be equivalent to selecting all 20 plots. No participant was shown the same lineup twice. Information about preferred pronouns, age group, education, and previous experience in visual experiments were also collected. A participant's submission was only included in the analysis if the data plot is identified for at least one attention check.

@@

With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation. Precise estimation of $\alpha$, is aided by evaluation of Rorschach lineups, where all plots are null plots. In a Rorschach, in theory all plots should be equally likely, but in practice some (irrelevant) visual elements may be more eye-catching than others. This is what $alpha$ captures, the capacity for extraneous features to distract the observer for a particular type of plot display.

@@

Let $\boldsymbol{X}$ denotes the $p + 1$ regressors with $n$ observations, $\boldsymbol{b} = (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y}$ denotes the OLS solution, $\boldsymbol{R} = \boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error. The residual vector

@@

$$KL = \frac{1}{2}\left(log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}
\text{diag}(\boldsymbol{R}\sigma^2))\right).$$
Since we assume $\sigma = 1$ for the heteroskedasticity model, the final form of the formula is

$$KL = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R})|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}))\right).$$

@@

```{r rej-rates, fig.pos="t!", fig.width=8, fig.height=5, out.width="100%", fig.cap="Changing the significance level of the conventional test will change the rejection rate. A: Percentage of conventional tests also rejected by the visual test: rejected (solid), not rejected (dashed). As the significance level is reduced, the percentage rejected by the visual test that has been rejected by the conventional test approaches 100. The percentage of tests not rejected by the conventional test also increases, as expected, but the percentage of these that are rejected by the visual test increases too. B: ROC curve shows that forcing the conventional test to not reject creates a discrepancy with the visual test. Many of the residual plots not rejected by the conventional test are rejected by the visual test. It isn't possible to vary the significance level of the conventional test to match the decisions made by the visual test."}
# compute rejection rate for different significance levels, holding visual test constant
library(ggpubr)

@@

# Experiment setup

>>>>

---
title: |
  Appendix to "A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol" published in the Journal of Computational and Graphical Statistics
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
  - name: Dianne Cook
    affil: a
    email: dicook@monash.edu
  - name: Emi Tanaka
    affil: a, b, c
    email: emi.tanaka@anu.edu.au
  - name: Susan VanderPlas
    affil: d
    email: susan.vanderplas@unl.edu
affiliation:
  - num: a
    address: |
      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
  - num: b
    address: |
      Biological Data Science Institute, Australian National University, Acton, ACT, Australia
  - num: c
    address: |
      Research School of Finance, Actuarial Studies and Statistics, Australian National University, Acton, ACT, Australia
  - num: d
    address: |
      Department of Statistics, University of Nebraska, Lincoln, Nebraska, USA
bibliography: ref.bib
header-includes: |
  \usepackage{lscape}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
  \doublespacing
output: rticles::tf_article
editor_options: 
  chunk_output_type: console
---

@@

@@

With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation. Precise estimation of $\alpha$, is aided by evaluation of Rorschach lineups, where all plots are null plots. In a Rorschach, in theory all plots should be equally likely, but in practice some (irrelevant) visual elements may be more eye-catching than others. This is what $\alpha$ captures, the capacity for extraneous features to distract the observer for a particular type of plot display.

@@

Let $\boldsymbol{X}$ denotes the $p + 1$ predictors with $n$ observations, $\boldsymbol{b} = (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y}$ denotes the OLS solution, $\boldsymbol{R} = \boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error. The residual vector

@@

$$KL = \frac{1}{2}\left(log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}
\text{diag}(\boldsymbol{R}\sigma^2))\right).$$

To compute the effect size for each lineup we simulate a sufficiently large number of samples from the same model with the number of observations $n$ fixed for each sample. We then compute the effect size for each sample and take the average as the final value. This ensures lineups constructed with the same experimental factors will share the same effect size.

@@

```{r rej-rates, fig.pos="t!", fig.width=8, fig.height=5, out.width="100%", fig.cap="Changing the significance level of the conventional test will change the rejection rate. A: Percentage of conventional tests also rejected by the visual test: rejected (solid), not rejected (dashed). As the significance level is reduced, the percentage rejected by the visual test that has been rejected by the conventional test approaches 100. The percentage of tests not rejected by the conventional test also increases, as expected, but the percentage of these that are rejected by the visual test increases too. B: ROC curve shows that forcing the conventional test to not reject creates a discrepancy with the visual test. Many of the residual plots not rejected by the conventional test are rejected by the visual test. It is not possible to vary the significance level of the conventional test to match the decisions made by the visual test."}
# compute rejection rate for different significance levels, holding visual test constant
library(ggpubr)

@@

# Additional details of experiment setup

## Non-linearity Model

## Heteroskedasticity Model

Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively. A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern. 
The parameter $b$ controls the variation in the standard deviation of the error across the support of the predictor. A larger value of $b$ generates a larger ratio between it is smallest and highest values, making the visual pattern more obvious. 
The sample size $n$ sharpens (or blurs) a pattern when it is larger (or smaller).

## Lineup allocation to participants

There are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ combinations of parameter values for the non-linearity model and heteroskedasticity models respectively. Three replications for each combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups, respectively.

Each lineup needs to be evaluated by at least five participants. From previous work, and additional pilot studies for this experiment, we decided to record evaluations from 20 lineups for each participant. Two of the 20 lineups with clear visual patterns were used as attention checks, to ensure quality data. Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ participants were needed to cover the experimental design for the data collection periods I and II, respectively. The factor levels and range of difficulty was assigned relatively equally among participants.

Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$, and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results. To construct a Rorschach lineup, the data is generated from a model with zero effect size, while the null data are generated using the residual rotation technique. This procedure differs from that of the canonical Rorschach lineup, where all 20 plots are generated directly from the null model, so the method suggested in @vanderplas2021statistical for typical lineups containing a data plot is used to estimate $\alpha$. The $3 \times 4 = 12$ treatment levels of the common factors, replicated three times results in 36 lineups. And 6 more evaluations on the 279 lineups with uniform fitted value distribution, results in at least $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants needed.

### 2024-03-24 appendix.tex
R2 submitted

\title{Appendix: A Plot is Worth a Thousand Tests: Assessing Residual
Diagnostics with the Lineup Protocol}

\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi
Tanaka$^{a}$, Susan VanderPlas$^{b}$}
\affil{$^{a}$Department of Econometrics and Business Statistics, Monash
University, Clayton, VIC, Australia; $^{b}$Department of Statistics,
University of Nebraska, Lincoln, Nebraska, USA}
}

\thanks{CONTACT Weihao
Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}, Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Emi
Tanaka. Email: \href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}, Susan
VanderPlas. Email: \href{mailto:susan.vanderplas@unl.edu}{\nolinkurl{susan.vanderplas@unl.edu}}}

@@

\hypertarget{additional-details-of-testing-procedures}{%
\section{Additional details of testing
procedures}\label{additional-details-of-testing-procedures}}

\hypertarget{statistical-significance}{%
\subsection{Statistical significance}\label{statistical-significance}}

@@

With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
estimated using maximum likelihood estimation. Precise estimation of
\(\alpha\), is aided by evaluation of Rorschach lineups, where all plots
are null plots. In a Rorschach, in theory all plots should be equally
likely, but in practice some (irrelevant) visual elements may be more
eye-catching than others. This is what \(alpha\) captures, the capacity
for extraneous features to distract the observer for a particular type
of plot display.

\hypertarget{effect-size-derivation}{%
\subsection{Effect size derivation}\label{effect-size-derivation}}

@@

Let \(\boldsymbol{X}\) denotes the \(p + 1\) regressors with \(n\)
observations,
\(\boldsymbol{b} = (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y}\)
denotes the OLS solution,
\(\boldsymbol{R} = \boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
denotes the residual operator, and let
\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})\)
denotes the error. The residual vector

@@

\[KL = \frac{1}{2}\left(log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}
\text{diag}(\boldsymbol{R}\sigma^2))\right).\] Since we assume
\(\sigma = 1\) for the heteroskedasticity model, the final form of the
formula is

\[KL = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R})|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}))\right).\]

\hypertarget{sensitivity-analysis-for-alpha}{%
\subsection{\texorpdfstring{Sensitivity analysis for
\(\alpha\)}{Sensitivity analysis for \textbackslash alpha}}\label{sensitivity-analysis-for-alpha}}

@@

\hypertarget{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}{%
\subsection{Effect of number of evaluations on the power of a visual
test}\label{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}}

@@

\hypertarget{power-of-a-reset-test-under-different-auxiliary-regression-formulas}{%
\subsection{Power of a RESET test under different auxiliary regression
formulas}\label{power-of-a-reset-test-under-different-auxiliary-regression-formulas}}

@@

\hypertarget{conventional-test-rejection-rate-for-varying-significance-levels}{%
\subsection{Conventional test rejection rate for varying significance
levels}\label{conventional-test-rejection-rate-for-varying-significance-levels}}

@@

\caption{Changing the significance level of the conventional test will change the rejection rate. A: Percentage of conventional tests also rejected by the visual test: rejected (solid), not rejected (dashed). As the significance level is reduced, the percentage rejected by the visual test that has been rejected by the conventional test approaches 100. The percentage of tests not rejected by the conventional test also increases, as expected, but the percentage of these that are rejected by the visual test increases too. B: ROC curve shows that forcing the conventional test to not reject creates a discrepancy with the visual test. Many of the residual plots not rejected by the conventional test are rejected by the visual test. It isn't possible to vary the significance level of the conventional test to match the decisions made by the visual test.}\label{fig:rej-rates}
\end{figure}

\hypertarget{experiment-setup}{%
\section{Experiment setup}\label{experiment-setup}}

\hypertarget{mapping-of-participants-to-experimental-factors}{%
\subsection{Mapping of participants to experimental
factors}\label{mapping-of-participants-to-experimental-factors}}

@@

\hypertarget{data-collection-process}{%
\subsection{Data collection process}\label{data-collection-process}}

@@

\hypertarget{analysis-of-results-relative-to-data-collection-process}{%
\section{Analysis of results relative to data collection
process}\label{analysis-of-results-relative-to-data-collection-process}}

\hypertarget{demographics}{%
\subsection{Demographics}\label{demographics}}

@@

\hypertarget{data-collection-periods}{%
\subsection{Data collection periods}\label{data-collection-periods}}

>>>>

\title{Appendix to ``A Plot is Worth a Thousand Tests: Assessing
Residual Diagnostics with the Lineup Protocol'' published in the Journal
of Computational and Graphical Statistics}

\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi Tanaka$^{a, b,
c}$, Susan VanderPlas$^{d}$}
\affil{$^{a}$Department of Econometrics and Business Statistics, Monash
University, Clayton, VIC, Australia; $^{b}$Biological Data Science
Institute, Australian National University, Acton, ACT,
Australia; $^{c}$Research School of Finance, Actuarial Studies and
Statistics, Australian National University, Acton, ACT,
Australia; $^{d}$Department of Statistics, University of Nebraska,
Lincoln, Nebraska, USA}
}

\thanks{CONTACT Weihao
Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}, Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Emi
Tanaka. Email: \href{mailto:emi.tanaka@anu.edu.au}{\nolinkurl{emi.tanaka@anu.edu.au}}, Susan
VanderPlas. Email: \href{mailto:susan.vanderplas@unl.edu}{\nolinkurl{susan.vanderplas@unl.edu}}}

@@

\section{Additional details of testing
procedures}\label{additional-details-of-testing-procedures}

\subsection{Statistical significance}\label{statistical-significance}

@@

With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
estimated using maximum likelihood estimation. Precise estimation of
\(\alpha\), is aided by evaluation of Rorschach lineups, where all plots
are null plots. In a Rorschach, in theory all plots should be equally
likely, but in practice some (irrelevant) visual elements may be more
eye-catching than others. This is what \(\alpha\) captures, the capacity
for extraneous features to distract the observer for a particular type
of plot display.

\subsection{Effect size derivation}\label{effect-size-derivation}

@@

Let \(\boldsymbol{X}\) denotes the \(p + 1\) predictors with \(n\)
observations,
\(\boldsymbol{b} = (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y}\)
denotes the OLS solution,
\(\boldsymbol{R} = \boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
denotes the residual operator, and let
\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})\)
denotes the error. The residual vector

@@

\[KL = \frac{1}{2}\left(log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}
\text{diag}(\boldsymbol{R}\sigma^2))\right).\]

To compute the effect size for each lineup we simulate a sufficiently
large number of samples from the same model with the number of
observations \(n\) fixed for each sample. We then compute the effect
size for each sample and take the average as the final value. This
ensures lineups constructed with the same experimental factors will
share the same effect size.

\subsection{\texorpdfstring{Sensitivity analysis for
\(\alpha\)}{Sensitivity analysis for \textbackslash alpha}}\label{sensitivity-analysis-for-alpha}

@@

\subsection{Effect of number of evaluations on the power of a visual
test}\label{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}

@@

\subsection{Power of a RESET test under different auxiliary regression
formulas}\label{power-of-a-reset-test-under-different-auxiliary-regression-formulas}

@@

\subsection{Conventional test rejection rate for varying significance
levels}\label{conventional-test-rejection-rate-for-varying-significance-levels}

@@

\caption{Changing the significance level of the conventional test will change the rejection rate. A: Percentage of conventional tests also rejected by the visual test: rejected (solid), not rejected (dashed). As the significance level is reduced, the percentage rejected by the visual test that has been rejected by the conventional test approaches 100. The percentage of tests not rejected by the conventional test also increases, as expected, but the percentage of these that are rejected by the visual test increases too. B: ROC curve shows that forcing the conventional test to not reject creates a discrepancy with the visual test. Many of the residual plots not rejected by the conventional test are rejected by the visual test. It is not possible to vary the significance level of the conventional test to match the decisions made by the visual test.}\label{fig:rej-rates}
\end{figure}

\section{Additional details of experiment
setup}\label{additional-details-of-experiment-setup}

\subsection{Non-linearity Model}\label{non-linearity-model}

The non-linearity model used in the experiment includes a non-linear
term \(\boldsymbol{z}\) constructed using Hermite polynomials on random
vector \(\boldsymbol{x}\) formulated as

\begin{align*} \label{eq:nonlinearity-model}
\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
\end{align*}

\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(\boldsymbol{1}_n\)
is a vector of ones of size \(n\), \(He_{j}(.)\) is the \(j\)th-order
probabilist's Hermite polynomials
\citetext{\citealp{hermite1864nouveau}; \citealp[originally
by][]{de1820theorie}},
\(\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)\), and
\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
the random vector to be \([-k, k]^n\) defined as

\begin{equation} \label{eq:scaling-function}
g(\boldsymbol{x}, k) = 2k\cdot\frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k, \quad \text{for} \quad k > 0. 
\end{equation}

\noindent where
\(\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i\),
\(\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i\) and \(x_i\) is
the \(i\)-th entry of \(\boldsymbol{x}\). The function \texttt{hermite}
from the R package \texttt{mpoly} \citep{mpoly} is used to simulate
\(\boldsymbol{z}_{raw}\) to generate Hermite polynomials.

The null regression model used to fit the realizations generated by the
above model is formulated as

\begin{equation} \label{eq:null-model}
\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

\noindent where
\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
Here \(\boldsymbol{z}\) is a higher order term of \(\boldsymbol{x}\) and
leaving it out in the null regression results in model misspecification.

\subsection{Heteroskedasticity Model}\label{heteroskedasticity-model}

The heteroskedasticity model used in the experiment is formulated as

\begin{align*} \label{eq:heter-model}
\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
\end{align*}

\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
the scaling function defined in Equation \ref{eq:scaling-function}. The
null regression model used to fit the realizations generated by the
above model is formulated exactly the same as Equation
\ref{eq:null-model}.

For \(b \neq 0\), the variance-covariance matrix of the error term
\(\boldsymbol{\varepsilon}\) is correlated with the predictor
\(\boldsymbol{x}\), which will lead to the presence of
heteroskedasticity.

Since \(\text{supp}(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\)
and \(1\) can generate ``left-triangle'', ``butterfly'' and
``right-triangle'' shapes. The term \((2 - |a|)\) maintains the
magnitude of residuals across different values of \(a\).

\subsection{Controlling the signal
strength}\label{controlling-the-signal-strength}

The three parameters \(n\), \(\sigma\) and \(b\) are important for
controlling the strength of the signal to generate lineups with a
variety difficulty levels. This will ensure that estimated power curves
will be smooth and continuous, and that participants are allocated a set
of lineups with a range of difficulty.

Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
\(b \in \{0.25, 1, 4, 16, 64\}\) are used in data collection periods I
and II respectively. A large value of \(\sigma\) will increase the
variation of the error of the non-linearity model and decrease the
visibility of the visual pattern. The parameter \(b\) controls the
variation in the standard deviation of the error across the support of
the predictor. A larger value of \(b\) generates a larger ratio between
it is smallest and highest values, making the visual pattern more
obvious. The sample size \(n\) sharpens (or blurs) a pattern when it is
larger (or smaller).

\subsection{Lineup allocation to
participants}\label{lineup-allocation-to-participants}

There are a total of \(4 \times 4 \times 3 \times 4 = 192\) and
\(3 \times 5 \times 3 \times 4 = 180\) combinations of parameter values
for the non-linearity model and heteroskedasticity models respectively.
Three replications for each combination results in
\(192 \times 3 = 576\) and \(180 \times 3 = 540\) lineups, respectively.

Each lineup needs to be evaluated by at least five participants. From
previous work, and additional pilot studies for this experiment, we
decided to record evaluations from 20 lineups for each participant. Two
of the 20 lineups with clear visual patterns were used as attention
checks, to ensure quality data. Thus, \(576 \times 5 / (20-2) = 160\)
and \(540 \times 5 / (20-2) = 150\) participants were needed to cover
the experimental design for the data collection periods I and II,
respectively. The factor levels and range of difficulty was assigned
relatively equally among participants.

Data collection period III was primarily to obtain Rorschach lineup
evaluations to estimate \(\alpha\), and also to obtain additional
evaluations of lineups made with uniform fitted value distribution to
ensure consistency in the results. To construct a Rorschach lineup, the
data is generated from a model with zero effect size, while the null
data are generated using the residual rotation technique. This procedure
differs from that of the canonical Rorschach lineup, where all 20 plots
are generated directly from the null model, so the method suggested in
\citet{vanderplas2021statistical} for typical lineups containing a data
plot is used to estimate \(\alpha\). The \(3 \times 4 = 12\) treatment
levels of the common factors, replicated three times results in 36
lineups. And 6 more evaluations on the 279 lineups with uniform fitted
value distribution, results in at least
\((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) participants
needed.

\subsection{Mapping of participants to experimental
factors}\label{mapping-of-participants-to-experimental-factors}

@@

\subsection{Data collection process}\label{data-collection-process}

@@

\section{Analysis of results relative to data collection
process}\label{analysis-of-results-relative-to-data-collection-process}

\subsection{Demographics}\label{demographics}

@@

\subsection{Data collection periods}\label{data-collection-periods}
